---
output: md_document
author: "Nancy Liu & Breanna Richards"
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = FALSE)
library(tidyverse)
library(lubridate)
library(ical)
library(kableExtra)
library(janitor)
library(reshape2)
library(data.table)
library(ggplot2)
library(ggthemes)
library(gganimate)
library(gifski)
library(survival)
library(randomForestSRC)
```

# Predicting Survival and Controlling for Biased with Random Survival Forests and Condtional Inference Forests Using Right-Censored Data



## What is Survival Analysis?

Survival 


## Review of Random Forests


## Random Survival Forests


## Conditional Inference Forests



## Methods



## Results


\textbf{PBC Data}

Primary sclerosing cholangitis is an autoimmune disease leading to destruction of the small bile ducts in the liver. Progression is slow but inexorable, eventually leading to cirrhosis and liver decompensation. 

This data is from the Mayo Clinic trial in PBC conducted between 1974 and 1984. A total of 424 PBC patients met eligibility criteria for the randomized placebo controlled trial of the drug D-penicillamine. This dataset tracks survival status until end up follow up period as well as contains many covariates collected during the clinical trial.

```{r}
# example
pbc
```


```{r}
pbc_rf <- rfsrc(Surv(years, status) ~ ., data = pbc,
 nsplit = 10,
 na.action = "na.impute")

```

